---
title: "The Generic Pattern among IRAP effects"
author: "Ian Hussey"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r, include=FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE,
                      fig.height = 3, 
                      fig.width = 6)
```

```{r}

# dependencies ----------------------------------------------------------------------------

library(tidyverse)
library(schoRsch)
library(knitr)
library(kableExtra)
library(ez)
library(schoRsch)
library(plotrix)
library(BayesFactor)
library(rsample)
library(broom)
library(purrr)
library(patchwork)

# function to round all numeric vars in a data frame
round_df <- function(df, n_digits = 3) {
  df %>% mutate_if(is.numeric, round, digits = n_digits)
}

# table format in output
options(knitr.table.format = "html") 

# disable scientific notation
options(scipen = 999) 

```

```{r}

# exclusions ----------------------------------------------------------------------------


# get data
data <- 
  read.csv("../data/processed/combined data.csv") %>%
  mutate(gender = ifelse(as.character(gender) == "", NA, as.character(gender)))


# exclusions ----------------------------------------------------------------------------


# find outliers
exclusions <- data %>%
  group_by(domain) %>%
  mutate(acc_outlier = ifelse(perc_acc < mean(perc_acc) - sd(perc_acc)*2.5, 1, 0),
         rt_outlier = ifelse(mean_rt > mean(mean_rt) + sd(mean_rt)*2.5, 1, 0)) %>%
  ungroup() %>%
  mutate(exclude = ifelse(rt_outlier == 1, 1,
                          ifelse(acc_outlier == 1, 1, 0))) %>%
  unite(study, c("domain", "researcher"), remove = FALSE)

# exclude outliers
data_wide <- exclusions %>%
  filter(exclude == 0) 

# reshape to long format for anova
data_long <- data_wide %>%
  gather(trialtype, D_score, c(D_tt1, D_tt2, D_tt3, D_tt4)) %>%
  dplyr::mutate(unique_id = as.factor(unique_id),
                trialtype = ifelse(trialtype == "D_tt1", "Category 1\npositive", 
                                   ifelse(trialtype == "D_tt2", "Category 1\nnegative", 
                                          ifelse(trialtype == "D_tt3", "Category 2\npositive", 
                                                 ifelse(trialtype == "D_tt4", "Category 2\nnegative", NA)))),
                trialtype = forcats::fct_relevel(as.factor(trialtype), 
                                                 "Category 1\npositive", 
                                                 "Category 1\nnegative", 
                                                 "Category 2\npositive", 
                                                 "Category 2\nnegative"))

data_long_known_stimuli <- data_long %>%
  filter(domain != "Non-words")

# write_csv(data_long_known_stimuli, "data_long_known_stimuli.csv")

data_long_unknown_stimuli <- data_long %>%
  mutate(stimuli = ifelse(domain == "Non-words", "Non-words", "Real words"),
         stimuli = as.factor(stimuli))


# excluded
n_total <- exclusions %>% summarize(n = n()) %>% as.integer()
n_excluded <- exclusions %>% filter(exclude == 1) %>% summarize(n = n()) %>% as.integer()
perc_excluded <- round(n_excluded/n_total*100, 0)



# sample descriptives ----------------------------------------------


# analytic sample after exclusions
summary_stats <- data_wide %>%
  dplyr::summarise(age_mean = round(mean(age, na.rm = TRUE), 1), 
                   age_sd = round(sd(age, na.rm = TRUE), 1), 
                   rt_mean = round(mean(mean_rt), 0),
                   rt_sd = round(sd(mean_rt), 0),
                   perc_acc_mean = round(mean(perc_acc)*100, 1),
                   perc_acc_sd = round(sd(perc_acc)*100, 1))


gender_summary <- data_wide %>%
  group_by(gender) %>%
  dplyr::summarise(n = n())

```

Data from k = `r exclusions %>% distinct(study) %>% summarize(n = n()) %>% as.integer()` studies across k = `r exclusions %>% distinct(domain) %>% summarize(n = n()) %>% as.integer()` domains were selected for inclusion (some data originally published in Drake et al., 2010, Hussey, Daly, & Barnes-Holmes 2015). This included a total of N = `r n_total` participants.

# Demographics

The analytic sample after exclusions included `r gender_summary %>% filter(gender == "Female") %>% select(n) %>% as.integer()` women, `r gender_summary %>% filter(gender == "Male") %>% select(n) %>% as.integer()` women, and `r gender_summary %>% filter(gender == "Other") %>% select(n) %>% as.integer()` identified using a third gender category (*M* = `r summary_stats$age_mean`, *SD* = `r summary_stats$age_sd`). `r gender_summary %>% filter(is.na(gender)) %>% select(n) %>% as.integer()` had missing demographics data. Accuracy and latency performances on the IRAP were typical of previous studies (*M*~RT~ = `r summary_stats$rt_mean`, *SD* = `r summary_stats$rt_sd`; *M*~accuracy~ = `r summary_stats$perc_acc_mean`, *SD* = `r summary_stats$perc_acc_sd`).

N participants by IRAP domain:

```{r}

data_wide %>%
  group_by(domain) %>%
  dplyr::summarise(n = n()) %>% 
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```

# Results

## Comparing domains

### Plot

```{r, fig.height=3, fig.width=5.5}

data_summary <- data_long_known_stimuli %>%
  group_by(domain, trialtype) %>%
  dplyr::rename(Domain = domain) %>%
  dplyr::summarise(mean_D = mean(D_score),
                   se_D = std.error(D_score)) %>%
  round_df(2)

p1 <-
  ggplot(data_summary, aes(x = trialtype, y = mean_D, colour = Domain)) +
  geom_linerange(aes(ymax = mean_D + (1.96*se_D),
                     ymin = mean_D + (-1.96*se_D)),
                 position = position_dodge(width = .5)) +
  geom_point(aes(shape = Domain),
             size = 2.5,
             position = position_dodge(width = .5)) +
  geom_line(aes(group = Domain),
            position = position_dodge(width = .5)) +
  geom_hline(yintercept=c(0), linetype="dotted") +
  ylab(expression(paste(italic("D"), " score"))) +
  xlab("Trial type") +
  theme_classic()

p1

```

### Descriptives

For the generic pattern (i.e., collapsed across all known word domains)

```{r}

data_summary_generic_pattern <- data_long_known_stimuli %>%
  group_by(trialtype) %>%
  dplyr::summarise(mean_D = mean(D_score),
                   se_D = std.error(D_score),
                   sd_D = sd(D_score)) %>%
  round_df(2)

data_summary_generic_pattern %>% 
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```

### Bayes Factors 

These use the equivalent of frequentist type II sum of squares model comparisons for the main effects. 

- BF for the evidence for domain is the BF(D_score ~ trial_type + domain + unique_id)/BF(D_score ~ trial_type + unique_id)
- BF for the evidence for trial type is the BF(D_score ~ trial_type + domain + unique_id)/BF(D_score ~ domain + unique_id)

And the equivalent of frequentist type III sum of squares model comparisons for the interaction effect.

- BF for the evidence for trial type is the BF(D_score ~ trial_type + domain + trial_type:domain + unique_id)/BF(D_score ~ trial_type + domain + unique_id)

```{r}


if (file.exists("models/fit_bf_anova.RData")) {
  
  load("models/fit_bf_anova.RData")

} else {

  fit_bf_anova <-
    anovaBF(formula = D_score ~ trialtype * domain + unique_id,
            data = data_long_known_stimuli,
            whichRandom = c("unique_id"),
            multicore = TRUE)
  
  save(fit_bf_anova, file = "models/fit_bf_anova.RData")
  
}

# BF inclusions (between multiple alternative models)
bf_inclusion_domain      <- fit_bf_anova[3] / fit_bf_anova[2]
bf_inclusion_trialtype   <- fit_bf_anova[3] / fit_bf_anova[1]
bf_inclusion_interaction <- fit_bf_anova[4] / fit_bf_anova[3]

```

- Domain BF10 = `r round(exp(bf_inclusion_domain@bayesFactor$bf), 4)`
- Trial type BF10 =  `r round(exp(bf_inclusion_trialtype@bayesFactor$bf), 2)`
- Interaction effect BF10 = `r round(exp(bf_inclusion_interaction@bayesFactor$bf), 4)`

### Frequentist 

```{r}

fit_anova_freq <- ez::ezANOVA(data     = data_long_known_stimuli,
                              dv       = D_score,
                              within   = trialtype,
                              between  = domain,
                              wid      = unique_id,
                              type     = 3,
                              detailed = TRUE)

# results
anova_out(fit_anova_freq, 
          etasq = "partial", 
          print = FALSE)$`--- FORMATTED RESULTS ------------------------------------` 

```

### Effect sizes

Comparison of frequentist effect sizes: eta squared, partial eta squared, and generalized eta squared.

Bootstrapping via case removal and percentile method with 1000 replications.

```{r}

if (file.exists("models/boot_fits.RData")) {
  
  load("models/boot_fits.RData")

} else {
  
  # create bootstraps using out of bag method. makes a df with values that are collapsed dfs.
  boots <- data_long_known_stimuli %>%
    # reshape so that participants are sampled between bootstraps
    select(unique_id, domain, trialtype, D_score) %>%
    spread(trialtype, D_score) %>%
    rename(tt1 = `Category 1\npositive`,
           tt2 = `Category 1\nnegative`,
           tt3 = `Category 2\npositive`,
           tt4 = `Category 2\nnegative`) %>%
    # create bootstraps
    bootstraps(times = 1000)
  
  # analysis workflow to be bootstrapped
  helper_function <- function(split) {
    
    fit_anova_freq <- analysis(split) %>%
      # undo reshaping
      gather(trialtype, D_score, c("tt1", "tt2", "tt3", "tt4")) %>%
      # fit anova
      ez::ezANOVA(data     = .,
                  dv       = D_score,
                  within   = trialtype,
                  between  = domain,
                  wid      = unique_id,
                  type     = 3,
                  detailed = TRUE)
    
    # eta squared and generalized eta squared
    results_etasq <- 
      fit_anova_freq$ANOVA %>%
      mutate(etasq = SSn / sum(SSn)) %>%
      select(Effect, etasq) %>%
      round_df(2)
    
    # partial eta squared
    results_petasq_and_getasq <- 
      anova_out(fit_anova_freq, 
                etasq = "partial", 
                print = FALSE)$`--- ANOVA RESULTS     ------------------------------------` %>%
      mutate(petasq = as.numeric(as.character(petasq)),
             getasq = as.numeric(as.character(getasq))) %>%
      select(Effect, petasq, getasq)
    
    # combine
    results_proportion_of_variance <- 
      left_join(results_etasq, results_petasq_and_getasq, by = "Effect")
    
    return(results_proportion_of_variance)
  }
  
  # apply to each bootstrap
  boot_fits <- boots %>% 
    mutate(fit = map(splits, helper_function)) %>%
    unnest(fit) 
  
  save(boot_fits, file = "models/boot_fits.RData")

}

boot_estimates <- boot_fits %>%
    # find CIs using percentile method
    group_by(Effect) %>%
    summarize(etasq_median    = quantile(etasq,  0.50),
              etasq_ci_lower  = quantile(etasq,  0.05),
              etasq_ci_upper  = quantile(etasq,  0.95),
              petasq_median   = quantile(petasq, 0.50),
              petasq_ci_lower = quantile(petasq, 0.05),
              petasq_ci_upper = quantile(petasq, 0.95),
              getasq_median   = quantile(getasq, 0.50),
              getasq_ci_lower = quantile(getasq, 0.05),
              getasq_ci_upper = quantile(getasq, 0.95))

boot_estimates %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```

NB 90% CIs as squared effect sizes are always positive!

## Comparing known and unknown words

### Plot

```{r, fig.height=3, fig.width=5.5}

# plot 1
data_means <- data_long_unknown_stimuli %>%
  group_by(stimuli, trialtype) %>%
  dplyr::rename(Stimuli = stimuli) %>%
  dplyr::summarise(mean_D = mean(D_score),
                   se_D = std.error(D_score)) %>%
  round_df(2)

p2 <-
  ggplot(data_means, aes(x = trialtype, y = mean_D, colour = Stimuli)) +
  geom_linerange(aes(ymax = mean_D + (1.96*se_D),
                    ymin = mean_D + (-1.96*se_D)),
                position = position_dodge(width = .2)) +
  geom_point(aes(shape = Stimuli),
             size = 2.5,
             position = position_dodge(width = .2)) +
  geom_line(aes(group = Stimuli),
            position = position_dodge(width = .2)) +
  geom_hline(yintercept=c(0), linetype="dotted") +
  ylab(expression(paste(italic("D"), " score"))) +
  xlab("Trial type") +
  theme_classic() +
  scale_color_viridis_d(begin = 0.3, end = 0.7)

p2

```

### Bayes Factors 

These use the equivalent of frequentist type II sum of squares model comparisons for the main effects. 

- BF for the evidence for domain is the BF(D_score ~ trial_type + stimuli + unique_id)/BF(D_score ~ trial_type + unique_id)
- BF for the evidence for trial type is the BF(D_score ~ trial_type + stimuli + unique_id)/BF(D_score ~ stimuli + unique_id)

And the equivalent of frequentist type III sum of squares model comparisons for the interaction effect.

- BF for the evidence for trial type is the BF(D_score ~ trial_type + stimuli + trial_type:stimuli + unique_id)/BF(D_score ~ trial_type + stimuli + unique_id)

```{r}

if (file.exists("models/fit_bf_anova_2.RData")) {
  
  load("models/fit_bf_anova_2.RData")
  
} else {
  
  fit_bf_anova_2 <-
    anovaBF(formula = D_score ~ trialtype * stimuli + unique_id,
            data = data_long_unknown_stimuli,
            whichRandom = c("unique_id"),
            multicore = TRUE)
  
  save(fit_bf_anova_2, file = "models/fit_bf_anova_2.RData")
  
}

# BF inclusions (between multiple alternative models)
bf_inclusion_stimuli     <- fit_bf_anova_2[3] / fit_bf_anova_2[1]
bf_inclusion_trialtype   <- fit_bf_anova_2[3] / fit_bf_anova_2[2]
bf_inclusion_interaction <- fit_bf_anova_2[4] / fit_bf_anova_2[3]

```

- Stimuli BF10 = `r round(exp(bf_inclusion_stimuli@bayesFactor$bf), 2)`
- Trial type BF10 =  `r round(exp(bf_inclusion_trialtype@bayesFactor$bf), 2)`
- Interaction effect BF10 = `r round(exp(bf_inclusion_interaction@bayesFactor$bf), 2)`

### Frequentist 

```{r}

fit_anova_freq_2 <- ez::ezANOVA(data     = data_long_unknown_stimuli,
                                dv       = D_score,
                                within   = trialtype,
                                between  = stimuli,
                                wid      = unique_id,
                                type     = 3,
                                detailed = TRUE)

# results
anova_out(fit_anova_freq_2, 
          etasq = "partial", 
          print = FALSE)$`--- FORMATTED RESULTS ------------------------------------` 

```

### Effect sizes

Comparison of frequentist effect sizes: eta squared, partial eta squared, and generalized eta squared.

Bootstrapping via case removal and percentile method with 1000 replications.

```{r}

if (file.exists("models/boot_fits_2.RData")) {
  
  load("models/boot_fits_2.RData")

} else {
  
  # create bootstraps using out of bag method. makes a df with values that are collapsed dfs.
  boots_2 <- data_long_unknown_stimuli %>%
    # reshape so that participants are sampled between bootstraps
    select(unique_id, stimuli, trialtype, D_score) %>%
    spread(trialtype, D_score) %>%
    rename(tt1 = `Category 1\npositive`,
           tt2 = `Category 1\nnegative`,
           tt3 = `Category 2\npositive`,
           tt4 = `Category 2\nnegative`) %>%
    # create bootstraps
    bootstraps(times = 1000)
  
  # analysis workflow to be bootstrapped
  helper_function_2 <- function(split) {
    
    fit_anova_freq_2 <- analysis(split) %>%
      # undo reshaping
      gather(trialtype, D_score, c("tt1", "tt2", "tt3", "tt4")) %>%
      # fit anova
      ez::ezANOVA(data     = .,
                  dv       = D_score,
                  within   = trialtype,
                  between  = stimuli,
                  wid      = unique_id,
                  type     = 3,
                  detailed = TRUE)
    
    # eta squared and generalized eta squared
    results_etasq_2 <- 
      fit_anova_freq_2$ANOVA %>%
      mutate(etasq = SSn / sum(SSn)) %>%
      select(Effect, etasq) %>%
      round_df(2)
    
    # partial eta squared
    results_petasq_and_getasq_2 <- 
      anova_out(fit_anova_freq_2, 
                etasq = "partial", 
                print = FALSE)$`--- ANOVA RESULTS     ------------------------------------` %>%
      mutate(petasq = as.numeric(as.character(petasq)),
             getasq = as.numeric(as.character(getasq))) %>%
      select(Effect, petasq, getasq)
    
    # combine
    results_proportion_of_variance_2 <- 
      left_join(results_etasq_2, results_petasq_and_getasq_2, by = "Effect")
    
    return(results_proportion_of_variance_2)
  }
  
  # apply to each bootstrap
  boot_fits_2 <- boots_2 %>% 
    mutate(fit = map(splits, helper_function_2)) %>%
    unnest(fit) 
  
  save(boot_fits_2, file = "models/boot_fits_2.RData")

}

boot_estimates_2 <- boot_fits_2 %>%
    # find CIs using percentile method
    group_by(Effect) %>%
    summarize(etasq_median    = quantile(etasq,  0.50),
              etasq_ci_lower  = quantile(etasq,  0.05),
              etasq_ci_upper  = quantile(etasq,  0.95),
              petasq_median   = quantile(petasq, 0.50),
              petasq_ci_lower = quantile(petasq, 0.05),
              petasq_ci_upper = quantile(petasq, 0.95),
              getasq_median   = quantile(getasq, 0.50),
              getasq_ci_lower = quantile(getasq, 0.05),
              getasq_ci_upper = quantile(getasq, 0.95))

boot_estimates_2 %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

```

NB 90% CIs as squared effect sizes are always positive!

# Combined plot for manuscript

```{r fig.height=6, fig.width=6}

plot_1 <- p1 + p2 + plot_layout(ncol = 1)
plot_1

write_rds(plot_1, "models/plot_1.rds")

```

# Recommendations

Valid and invalid inferences using a hypothetical data for a commonly employed design within IRAP research (i.e., comparisons between two known groups).  

```{r fig.height=3, fig.width=6}

# data
data <- data.frame(Group = c("Intervention","Control","Intervention","Control",
                             "Intervention","Control","Intervention","Control"),
                   D_score = c(0.4, 0.3, 0.16, 0.13, -0.21, -0.15, -0.13, 0.15),
                   trial_type = c("White people\npositive",
                                  "White people\npositive",
                                  "White people\nnegative",
                                  "White people\nnegative",
                                  "Black people\npositive",
                                  "Black people\npositive",
                                  "Black people\nnegative",
                                  "Black people\nnegative"),
                   SE = c(0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05, 0.05)) %>%
  dplyr::mutate(trial_type = forcats::fct_relevel(as.factor(trial_type), 
                                                  "White people\npositive",
                                                  "White people\nnegative",
                                                  "Black people\npositive",
                                                  "Black people\nnegative"))

# plot
plot_2 <-
  ggplot(data, aes(x = trial_type, y = D_score, colour = Group)) +
  geom_linerange(aes(ymax = D_score + (1.96*SE),
                     ymin = D_score + (-1.96*SE)),
                 position = position_dodge(width = .3)) +
  geom_point(aes(shape = Group),
             size = 2.5,
             position = position_dodge(width = .3)) +
  geom_line(aes(group = Group),
            position = position_dodge(width = .3)) +
  geom_hline(yintercept = c(0), linetype = "dotted") +
  ylab(expression(paste(italic("D"), " score"))) +
  xlab("Trial type") +
  theme_classic() + 
  geom_line(data = data.frame(x = c(1, 1), y = c(0.02, 0.28)),
            aes(x = x, y = y), 
            colour = "black",
            linetype = "dashed",
            arrow = arrow(length = unit(0.20,"cm"), 
                          ends="both", 
                          type = "closed")) +
  annotate("label", x = 1, y = 0.15, label = "(a) Invalid") +
  geom_line(data = data.frame(x = c(2.96, 3.91), y = c(-0.11, 0.17)),
            aes(x = x, y = y), 
            colour = "black",
            linetype = "dashed",
            arrow = arrow(length = unit(0.20,"cm"), 
                          ends="both", 
                          type = "closed")) +
  annotate("label", x = 3.5, y = 0.10, label = "(b) Invalid") +
  geom_line(data = data.frame(x = c(4, 4), y = c(-0.111, 0.13)),
            aes(x = x, y = y), 
            colour = "black",
            linetype = "dashed",
            arrow = arrow(length = unit(0.20,"cm"), 
                          ends="both", 
                          type = "closed")) +
  annotate("label", x = 4, y = 0, label = "(c) Valid") +
  scale_color_viridis_d(begin = 0.3, end = 0.7)

plot_2

write_rds(plot_2, "models/plot_2.rds")

```



