---
title: "Tidy trial level data for the nonwords IRAP"
author: "Ian Hussey"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r, include=FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)
```

```{r}

# dependencies
library(tidyverse)

```

# Nonwords

*Necessary data corrections:*

The IRAP setup files accidentally swapped the target stimuli across conditions as well as the rules. In order to remedy this, the trial-types for all IRAP condition 2 (arbitrary IRAP only) should be reordered from 1234 to 3412. Then, TT1 and TT2 is always CUG, and TT3 and TT4 is always VEC.

Setup used in experiment:
IRAP 1
Target 1: CUG
Target 2: VEC	
CUG is positive and VEC is negative
CUG is negative and VEC is positive

IRAP 2
Target 1: VEC	[different to condition 1]
Target 2: CUG	[different to condition 1]
VEC is positive and CUG is negative
VEC is negative and CUG is positive


So, to make the two conditions produce data with equivalent interpretations, in condition 2 swap trialType 3,4 with 1,2, and swap currentRule 1 with 2.

after trial type re-labellings (both conditions):
tt1 = cug pos
tt2 = cug neg
tt3 = vec pos
tt4 = ved neg

```{r}

nonwords_condition_1 <- 
  read_csv("raw/nonwords condition 1 trimmed headers.csv") %>%
  mutate(condition = 1)

nonwords_condition_2 <- 
  read_csv("raw/nonwords condition 2 trimmed headers.csv") %>%
  mutate(condition = 2,
         trialType = case_when(trialType == 1 ~ 5,
                               trialType == 2 ~ 6,
                               trialType == 3 ~ 1,
                               trialType == 4 ~ 2)) %>%
  mutate(trialType = case_when(trialType == 5 ~ 3,
                               trialType == 6 ~ 4,
                               TRUE ~ trialType)) %>%
  mutate(currentRule = case_when(currentRule == 1 ~ 3,
                                  currentRule == 2 ~ 1)) %>%
  mutate(currentRule = case_when(currentRule == 3 ~ 2,
                                 currentRule == 1 ~ 1))
  
data_nonwords <- 
  bind_rows(nonwords_condition_1,
            nonwords_condition_2) %>%
  filter(isTestBlock == 1) %>%
  mutate(timepoint = "baseline", 
         domain = "Non-words",
         unique_id = paste(domain, participantCode)) %>%
  select(unique_id, 
         timepoint,
         domain,
         gender = participantGender,
         age = participantAge, 
         block_order = startingRule, 
         block_type = currentRule, 
         trial_type = trialType, 
         rt = timeToCorrectResponse,
         accuracy) %>%
  mutate(block_type = dplyr::recode(block_type,
                                    `1` = "con",
                                    `2` = "incon"),
         block_order = dplyr::recode(block_order,
                                     `1` = "con",
                                     `2` = "incon"),
         trial_type = dplyr::recode(trial_type,
                                    `1` = "tt1",
                                    `2` = "tt2",
                                    `3` = "tt3",
                                    `4` = "tt4")) %>%
  group_by(unique_id, timepoint) %>%
  mutate(mean_rt = round(mean(rt, na.rm = TRUE), 0)) %>%
  ungroup()

write_csv(data_nonwords, "data_trial_level_nonwords.csv")

```
